---
title: "Module 1: R basics"
author: "Jake Ferguson"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, "Title.css"]
    lib_dir: libs
    includes:
      after_body: logo.html
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
header-includes :
  - \usepackage {amsmath}
  - \usepackage{mathrsfs}
---
class: inverse

```{r, echo=F, message=F, warning=F, eval=F}
#setting up my ggplot defaults. Update with your own preferences
library(ggplot2)#plotting functions
library(ggthemes) #more themes!
library(wesanderson)
library(RColorBrewer)

theme_set(theme_tufte()) # a theme I like.
theme_update(plot.title = element_text(hjust = 0.5), 
             axis.line.x = element_line(color="black", size = 1),
             axis.line.y = element_line(color="black", size = 1),
             text=element_text(size=20),
             axis.text=element_text(size=15)) #center all titles and and axis lines
```

# Module goals

<br>

0. Navigate rstudio cloud

1. Understand how data is stored and represented in R

2. Read in datasets from files

3. Run some basic mathematical and statistical operations on the data 

4. View and summarize the datasets

---
# R, RStudio, and RStudio Cloud


- "R is a free software environment for statistical computing and graphics". This is the software that actually does the work.

  - download here: [https://cran.r-project.org/](https://cran.r-project.org/)
  
- "RStudio is an integrated development environment (IDE) for R, a programming language for statistical computing and graphics." This software is for interacting with R. 

  - download here: [https://rstudio.com/](https://rstudio.com/). You must install R first.

- "RStudio Cloud is a hosted version of RStudio in the cloud that makes it easy for professionals, hobbyists, trainers, teachers and students to do, share, teach and learn data science using R."

  - No download necessary: [https://rstudio.cloud/](https://rstudio.cloud/)

#### We will use RStudio Cloud exclusively in this course, but all work can reproduced in R and RStudio.

---
# Navigating `rstudio.cloud`


---
# Assigning values in R

The arrow <- is an assignment operation. 

The right-hand side is the value you assign

The left-hand side is the variable you assign a value to.


```{r}
x <- 3.14 #x is a number
y <- "CWS" #y is a character string
```

Anytime we refer to `x` and `y` in the future these values will be available.
```{r}
x
y
```


---
# Math operations

```{r}
x + 3

x*10

x/2

```

but not on characters
```{r, error=T}
y*2
```

Note that errors in R can be difficult to interpret. Cut and paste into **google** to decipher!
---
# Collections of values

A vector is a collection of values, `c()` combines values into a vector.

```{r}
w <- numeric(10) #A vector of 0's
x <- c(1, 2, 3, 4, 5, 6)  #this is a numeric vector
y <- c('yes', 'no', 'yes', 'no', 'yes', 'no') #this is a vector of characters 
```

we can do the math operations on numeric vectors

```{r}
w*1
x*3
x*x
```

these operations go elementwise, recycling values where appropriate.


---
# Functions

A function takes an argument does some stuff to it, then provides some output. 

```{r}
length(x) #returns the length of  vector
print(x) #function to print an object
mean(x) #calculate the mean of a vector
log(x) #calcuate the natural log of a vector
```

---
# Functions for creating vectors

R has some commands to create common vectors quickly that we will use.

```{r}
1:6 #create a sequence of integers with :
seq(from=1, to=6, by=0.5) #more complex sequences
rep(3.14, times=10) #repeat a value
```

---
class: inverse, middle

# Excercise 1A

## variables and vectors

---
# Data frames

We can collect vectors into a data frame. These vectors can be of different types.

```{r}
choice.dat <- data.frame(Number=x, Choice=y)
choice.dat
```

---
# Extracting and subsetting dataframes

We can access specific rows and columns of dataframes

```{r}
choice.dat[3:5,]

choice.dat[,2]

choice.dat[3:5,2]
```

or access columns by name with the `$` operator.

```{r}
choice.dat$Choice
```

---
# Reading in data from a file

Data files are often in the `csv` format (comma-seperated values). We'll learn later about excel files
```{r, eval=F}
bass.dat <- read.csv(file="BassGrowthOto.csv")
```

```{r, echo=F}
bass.dat <- read.csv(file="../Data/BassGrowthOto.csv")
```

The argument file is a string that points to the file you want to load. The above command assumed that the file was in the current working directory (where R loads up). This will always be the case on RStudio Cloud.

If the data file is in a different location, as it often is if you are working on RStudio on your computer, you tell R the filepath:
```{r,eval=F}
bass.dat <- read.csv(file="C::\\Users\Jake\BassGrowthOto.csv") #A windows path

bass.dat <- read.csv(file="/Users/Jake/BassGrowthOto.csv") #A mac or linux path

```

---
# Looking at your dataframe

Print the first 6 lines of the dataset (printing a large dataframe can freeze your machine)
```{r}
head(bass.dat)
```

Structure of the dataframe
```{r}
str(bass.dat)
```

---
class: inverse, middle

# Excercise 1B

## Reading in data

---
# Creating new variables

It is often useful to create new variables ...
