---
title: 'Module 5 worksheet'
output: html_notebook
---

```{r}
library(ggplot2) #load ggplot
library(ggthemes) #load themes package for ggplot
theme_set(theme_tufte()) # a theme I like.

#set some options that I like
theme_update(plot.title = element_text(hjust = 0.5), 
             axis.line.x = element_line(color="black", size = 1),
             axis.line.y = element_line(color="black", size = 1),
             text=element_text(size=20),
             axis.text=element_text(size=15)) #center all titles and and axis lines
```

# Exercise 7A
-----
## Fit a count model

Dodd and Silvertown [1999](https://doi.org/10.1086/303238) explored the relationships between size , age, competition, on the number of cones produced by fir (*Abies balsamea*). Interestingly, some of these populations experience wave-like die offs that lead to bands of dead trees in the forest (see image [here](https://www.conifers.org/pi/ab/balsamea13.jpg)). These waves are thought to be driven by chronic wind-stress. Today we will explore how fir fecundity is related to tree size, and population that have and not had wave die-offs.

- `DBH`: diameter in m at breast height (1.4 m)
- `fecundity`: number of cone rachises produced per year
- `pop`: whether the tree comes from a population that experiences non-wave or wave mortality patterns


1. Read in Firs.csv and plot the dataset, illustrating the effect of both `DBH` and population type (`pop`) on fecundity. What kind of distribution should we consider using to model fecundity? Based on this figure do you think the relationship between fecundity and DBH differs between populations?


```{r}
fir.dat <- read.csv(file="../Data/Firs.csv")
head(fir.dat)

ggplot(data=fir.dat, aes(x=DBH, y=fecundity, col=pop)) + 
  geom_point()
```

2. Now fit a Poisson regression to `fecundity` using `pop` and `DBH` as predictor variables. Report model estimates and confidence intervals, and plot the model fits using the function `plot_model`.

```{r}
firs.pois <- glm(fecundity ~ DBH + pop, data=fir.dat, family=poisson)
confint(firs.pois)
summary(firs.pois)
```
### Look, everything is supersignficant! This means we are either great scientists or our model has some problems. Guess which one...

###############
# Exercise 7B #
###############

## Negative binomial regression

Repeat the analysis in 7A but use a negative binomial model. Use AIC to determine which fit is better. Do the parameter estimates or confidence intervals differ much between the results in 7A and 7B?

```{r}
library(MASS)
library(MuMIn)
firs.nb <- glm.nb(fecundity ~ DBH + pop, data=fir.dat)
summary(firs.nb)
confint(firs.nb)
model.sel(list(firs.nb, firs.pois))
```
### We see some big differences when using the negative binomial (which is more strongly supported by AICc). The coefficeint  for the popwave predictor changes nad the parameter confidence interval widens by an order of magnitude.

###############
# Exercise 7C #
###############

## Modeling categorical data

Let's look at the silversword data from the lecture.

1. Load the dataset `silversword.csv`.
2. Build a survival model using the predictors `Pop_size`, `Precip`, `Elev`, and `veg_cover`. 
3. Increases in which predictors are associated with increases in survival? How about decreases in survival? Is there evidence of density dependence in the survival rate?

```{r}
silver.dat <- read.csv(file="../Data/Silversword.csv")
ggplot(data=silver.dat) +
  geom_point(mapping=aes(x=Elev, y=Survive/Pop_size)) + xlab("Elevation") + ylab("Survival probability")

full.mod <- glm(cbind(Survive, Die) ~ Elev + Precip + veg_cover + Pop_size, family=binomial, data=silver.dat)
library(car)
vif(full.mod)
summary(full.mod)
```

### We see that most of these effects are statistically significant. No obvious indications of problems such as collinearity or overdispersion. 


###############
# Exercise 7D #
###############

## Fit an overdispersed categorical model

1. Use the `VGAM` package to fit the model you used above to fit the same mean function as in 7C but using a betabinomial model. 
2. Compare the AIC values between the two model and compare their parameter estimates, are there systematic differences in the estimates?   

```{r}
library(VGAM) #contains beta-binomial regression
model.bb <- vglm(cbind(Survive, Die) ~ Elev + Precip + veg_cover + Pop_size, data=silver.dat, family=betabinomial)

coef(full.mod)
summary(model.bb)

```

### There are some slight differences in the estimates and precip has become non-signficant but doesn't appear to be anything to be concerned about in the original analysis.

