---
title: 'Module 4 worksheet'
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(ggplot2) #load ggplot
library(ggthemes) #load themes package for ggplot
theme_set(theme_tufte()) # a theme I like.

#set some options that I like
theme_update(plot.title = element_text(hjust = 0.5), 
             axis.line.x = element_line(color="black", size = 1),
             axis.line.y = element_line(color="black", size = 1),
             text=element_text(size=20),
             axis.text=element_text(size=15)) #center all titles and and axis lines
```

###############
# Exercise 4A #
###############

## Estimating means and standard deviations

Estimate the mean (`mean()`) and standard deviation (`sd()`) for the following data vectors:
```{r}
x <- 0:10
y <- rep(5, 10)
z <- sample(0:10, size=10, replace=TRUE) 
```

Rerun the analysis a couple of times do the estimates from `z` change or stay the same? Does this make sense?


###############
# Exercise 4B #
###############

## Calculating z-scores, and p-values 

1. Read in the dataset `ButterflyFishDiffs.csv`.
2. Calculate the standard error of the mean.
3. Claculate the confidence interval for the estimate of the mean.
4. Calculate the z-score for a hypothesis test with $H_0=0$ and corresponding p-value. 
5. State your conclusions.

```{r}
#code here
butt.dat <- read.csv(file="../Data/ButterflyFishDiffs.csv")
xbar <- mean(butt.dat$Difference)
se <- sd(butt.dat$Difference)/sqrt(16)

z.score <- xbar/se
p.val <- 2*pnorm(q=z.score, lower.tail=F)
CI.z <- xbar + c(-1,1)*qnorm(0.975)*se

CI.z
p.val
```


###############
# Exercise 4C #
###############

## Calculating Wald confidence intervals 

Using the dataset above, use the t-distribution to estimate the confidence intervals. Explain why these intervals are wider than the intervals using the normal distribution. Get the p-value as well. How does it compare to the p-value from the z-test in part B? How does this p-value compare to the value you got from the randomization test in Module 3?

```{r}
p.val <-2*pt(q=z.score, df=16-1, lower.tail=F)
CI.t <- xbar + c(-1,1)*qt(0.975, df=15)*se
CI.t
p.val

```


###############
# Exercise 4D #
###############

## One- and two-sample t-tests

For the following datasets determine if you should conduct a one- or two-sample test. Then conduct the appropriate test using `t.test()`. Note to conduct a paired t-test using the formula you can use `t.test(response ~ predictor, paired=TRUE)`.

1. The butterflyfish dataset: `ButterflyFish.csv`. One column is the male the second column is the female in a mated pair. 

2. Chen et al. (2011) studied recent range shifts in 31 taxa. This dataset (`RangeShift.csv`) has changes in highest detected elevation for these species from the late 1900's to the early 2000's. Postive numbers indicate upward shifts in elevation and negative numbers indicate shifts to lower elevations. Measurements are in meters.

3. In many species males are more likely to attract females if the males have high testosterone, one hypothesis is that the cost associated with high testosterone is a weakend ability to fight disease. Hasselquist et al. (1999) experimentally increased the amount of testosterone in 13 male red-winged blackbirds and measured their ability to produce antibodies to an antigen before and after implantation (`Blackbird.csv`). 

4. The neurodegenerative disease *Spinocerebellar Ataxia* type 1  is a disease marked by the gradual loss of motor skills and culminates in early death. Fryer et al. (2011) investigated the benefits of exercise in affecting the lifespan in a mouse model (`SpinocerebellarAtaxia.csv`).

```{r}
butt.dat <- read.csv(file="../Data/ButterflyFish.csv")
t.test(Length ~ Sex, data=butt.dat, paired=TRUE)

range.dat <- read.csv(file="../Data/RangeShift.csv")
t.test(x=range.dat$elevationalRangeShift)

bird.dat <- read.csv(file="../Data/Blackbird.csv")
t.test(Antibody ~ Treatment, data=bird.dat, paired=TRUE)

mouse.dat <- read.csv(file="../Data/SpinocerebellarAtaxia.csv")
t.test(lifespan ~ treatment, data=mouse.dat, paired=TRUE)
```

